Individual and Differential Harm in Redistricting

### 2025-08-25 paper/harm.qmd
edits from checking r1 comments

To allow for such averaging, we replace the fixed election $v$, counterfactual districting plan $\tilde q$, and individual voter $i$ with random variables $V\in\Elecs$, $\tilde Q\in\Plans$, and $I\in\Voters$.
The random variable $V$ can be conceptualized as ranging over possible future elections---one realization could correspond to a heavily Democratic-favoring midterm year, while another could involve incumbents under-performing particularly badly.
The random variable $\tilde Q$ captures the range of potential counterfactual districting plans.
Rarely would an analyst have a single counterfactual plan in mind: different realizations of $\tilde Q$ correspond to different districting plans that are permissible under the relevant districting criteria.
Finally, $I$ is defined to be a voter taken uniformly at random from $\Voters$, with $I\indep V,\tilde Q$.

@@

Since none of the existing measures, other than partisan dislocation, are centered on individual voters, they offer far less granular insights than the proposed harm framework, even in the case of studying partisan gerrymandering.
For example, differential harm can be estimated at the precinct level and then plotted on a map to show not just *which party* is harmed by a gerrymander, but *where* this harm occurs.

@@

The partisan dislocation measure [@defoeubarodd21] shows that the Democratic plan is typical ($p=0.24$) while the Republican plan is an outlier ($p=0.01$).
We use the vote-weighted measure of the average dislocation to produce a statewide summary.
We find that the sign is incorrect for the Democratic plan, which has a positive sign.
As the sign is positive for the simulations as well, this difference may be related to the fact that this measure uses the $k$ nearest neighbors, which is not uniform over the voters in the state.
Implicitly, voters in dense, urban areas will be counterd at a higher rate than those in rural areas.

>>>>

To allow for such averaging, we replace the fixed election $v$, counterfactual districting plan $\tilde q$, and individual voter $i$ with random variables $V\in\Elecs$, $\tilde Q\in\Plans$, and $I\in\Voters$.
The random variable $V$ can be conceptualized as ranging over possible future elections---one realization could correspond to a heavily Democratic-favoring midterm year, while another could involve incumbents under-performing particularly badly.
The random variable $\tilde Q$ captures the range of potential counterfactual districting plans.
Rarely would an analyst have a single counterfactual plan in mind: different realizations of $\tilde Q$ correspond to different districting plans that are permissible under the relevant districting criteria.
Finally, $I$ is defined to be a voter taken uniformly at random from $\Voters$, with $I$ independent of $V,\tilde Q$.

@@

Since none of the existing measures, except partisan dislocation, are centered on individual voters, they offer far less granular insights than the proposed harm framework, even in the case of studying partisan gerrymandering.
For example, differential harm can be estimated at the precinct level and then plotted on a map to show not just *which party* is harmed by a gerrymander, but *where* this harm occurs.

@@

The partisan dislocation measure [@defoeubarodd21] shows that the Democratic plan is typical ($p=0.24$) while the Republican plan is an outlier ($p=0.01$).
We use the vote-weighted measure of the average dislocation to produce a statewide summary.
We find that the sign is incorrect for the Democratic plan, which has a positive sign.
As the sign is positive for the simulations as well, this difference may be related to the fact that this measure uses the $k$ nearest neighbors, which is not uniform over the voters in the state.
Implicitly, voters in dense, urban areas will be counted at a higher rate than those in rural areas.
